---
title: "Descriptive Statistics"
output:
  html_document:
    df_print: paged
  pdf_document:
    keep_tex: yes
---

# Frequency distribution

## Discrete variables
Grades from two students:
```{r}
A <- factor(c(17, 16, 16, 16, 15, 16, 18, 17, 16, 16), levels = 14:18)
B <- factor(c(16, 15, 14, 17, 14, 17, 14, 14, 14, 14), levels = 14:18)
```

Absolute frequencies:
```{r}
table(A)
```

```{r}
prop.table(table(A))
```

```{r}
table(B)
```

```{r}
prop.table(table(B))
```

Bars plots:
```{r}
#png('010_010_distribution_discrete_barplot_A.png',
#   width = 6, height = 3, units = 'in', res = 512)

barplot(table(A), main = "A")

#dev.off()
```

```{r}
#png('010_010_distribution_discrete_barplot_B.png',
#    width = 6, height = 3, units = 'in', res = 512)

barplot(table(B), main = "B")

#dev.off()
```

## Continuous variables

```{r}
head(islands)
```

```{r}
ht <- hist(islands)
```

```{r}
distrib <- with(ht, data.frame(i = seq_along(counts),
                      from = head(breaks, n = -1),
                      to = tail(breaks, n = -1),
                      n_i = counts,
                      f_i = counts / sum(counts)))
distrib
```

```{r}
#png('010_020_distribution_continuous_equal_sizes.png',
#    width = 8, height = 5, units = 'in', res = 512)
#par(mar = c(4.2, 4.2, 2.1, 2.1))
hist(islands, xlab = 'square miles', main = '', col = 'gray')
#dev.off()
```

```{r}
# png('010_030_distribution_continuous_unequal_sizes.png',
#     width = 8, height = 5, units = 'in', res = 512)
# par(mar = c(4.2, 4.2, 2.1, 2.1))

ht2 <- hist(islands, breaks = c(0, 50, 100, 1000, 10000, 20000),
            main = '', xlab = 'square miles')

# dev.off()

distrib <- with(ht2, data.frame(i = seq_along(counts),
                               class_i = sprintf('(%d, %d]', head(breaks, n = -1), tail(breaks, n = -1)),
                               n_i = counts,
                               f_i = counts / sum(counts),
                               h_i = density))
distrib
```

```{r}
# png('010_030_distribution_continuous_unequal_sizes_zoom.png',
#     width = 8, height = 5, units = 'in', res = 512)
# par(mar = c(4.2, 4.2, 2.1, 2.1))

ht2 <- hist(islands, breaks = c(0, 50, 100, 1000, 10000, 20000),
            main = '', xlab = 'square miles',
            col = 'gray',
            xlim = c(0, 2000))

# dev.off()
```

## AJO - test1
```{r}
#trick to remove the levels (change from categrorical variable to vector)
A
as.character(A)
as.numeric(as.character(A))
```


## Empirical Cumulative Distribution Function

```{r}
A <- as.numeric(as.character(A))
B <- as.numeric(as.character(B))
Fn_A <- ecdf(A)
Fn_B <- ecdf(B)

# png('010_040_010_ecdf_students_A.png',
#     width = 6, height = 6, units = 'in', res = 512)

plot(Fn_A)

# dev.off()

# png('010_040_010_ecdf_students_B.png',
#     width = 6, height = 6, units = 'in', res = 512)

plot(Fn_B)

# dev.off()
```

```{r}
Fn_islands <- ecdf(islands)
#png('010_040_020_ecdf_islands.png',
#    width = 6, height = 6, units = 'in', res = 512)
plot(Fn_islands)
#dev.off()
```

## Quantile function

```{r}
quantile(A, probs = c(0.1, 0.5))
```

# Positional indexes

## AJO- side example
```{r}
X <- rnorm(50,mean=168,sd=0.2)

hist(X)
xm <- mean(X)
xm
#one wrong value (=outlier)
X[50]<-42
hist(X)
xm <- mean(X)
xm
#mean is not ok now.
xmed<-median(X)
xmed
#median is more robust to outliers.
```



## mean
```{r}
mean(islands)
```

## median
```{r}
median(islands)
```

## quartiles
```{r}
quantile(islands, c(0.25, 0.5, 0.75))
```

50% of the islands have surface area between 20.5 and 183.25 squared miles.

## boxplot
```{r}
# png('010_040_030_boxplot_islands.png',
#     width = 6, height = 6, units = 'in', res = 512)
boxplot(islands, log = "y")
# dev.off()
```

## variance
```{r}
var(islands)
sd(islands)
```

## coefficient of variation
```{r}
sd(islands) / mean(islands)
```

## interquartile range

```{r}
IQR(islands)
```

# Multivariate distributions

```{r}
library(car)

data(SLID)
head(SLID)
```



```{r}
# png('010_050_010_scatterplot.png',
#     width = 8, height = 6, units = 'in', res = 512)
plot(wages ~ education, pch = 16, cex = 0.4, data = SLID)
# dev.off()
```

```{r}
#png('010_050_020_boxplot.png',
#    width = 6, height = 6, units = 'in', res = 512)
plot(wages ~ sex, data = SLID)
# dev.off()
```

```{r}
library(ggplot2)

# png('010_050_030_histograms.png',
#     width = 8, height = 6, units = 'in', res = 512)
ggplot(SLID, aes(wages)) +
  geom_histogram() +
  facet_grid(sex ~ .)
# dev.off()
```

```{r}
# png('010_050_040_boxplot.png',
#     width = 6, height = 6, units = 'in', res = 512)
plot(wages ~ language, data = SLID)
# dev.off()
```

## Contingency tables
```{r}
with(SLID, table(sex, language))
```

## Linear Correlation Coefficient

```{r}
with(SLID, cor(education, wages))
```

```{r}
with(SLID, cor(education, wages, use = 'pairwise.complete.obs'))
```

```{r}
# png('010_050_050_scatterplot_lowess.png',
#     width = 8, height = 6, units = 'in', res = 512)
plot(wages ~ education, pch = 16, cex = 0.4, data = SLID)
lines(with(na.omit(SLID), lowess(education, wages)), col = 'blue', lwd = 2)
# dev.off()
```
