---
title: "Hurricanes Gender"
output:
  html_document:
    df_print: paged
---

Data from the study:

Jung, Kiju, Sharon Shavitt, Madhu Viswanathan, and Joseph M. Hilbe.
"Female hurricanes are deadlier than male hurricanes." Proceedings of the National Academy of Sciences 111, no. 24 (2014): 8782-8787.

Load libraries:
```{r}
library(tidyverse)
library(ggthemes) ## color blind friendly palette
theme_set(theme_bw())
```

Import the data in R:
```{r}
hur_raw <- read_csv('hurricanes.csv',
                col_types = 'cdcd')
hur_raw
```

```{r}
hur <- rename(hur_raw, name = Hurricane,
                       year = Year,
                       gender = `Gender of Name`,
                       deaths = `Number of Deaths`)
hur
```

```{r}
summary(hur)
```

Let's check there are no surprises:
```{r}
unique(hur$gender)
```


```{r}
hur <- mutate(hur, gender = factor(gender, levels = c('Female', 'Male')))
```

```{r}
summary(hur)
```

A cautionary note: please, DO NOT USE 'attach'/'detach'!!!

# Distribution

```{r}
hist(hur$deaths)
```

Boxplots:
```{r}
boxplot(deaths ~ gender, data = hur)
```

Those two extreme events might drive all our conclusions: let's put them aside for a moment:
```{r}
arrange(hur, desc(deaths))
```


```{r}
hur1 <- subset(hur, deaths < 150)
```

Let's redo the summaries:
```{r}
summary(hur1)
```

and boxplots:
```{r}
boxplot(deaths ~ gender, data = hur1)
```

We can also plot the histograms:
```{r}
ggplot(hur1, aes(x = deaths)) +
  geom_histogram() +
  facet_grid(gender ~ .)
```


Few numerical summaries:
```{r}
hur1 %>%
  group_by(gender) %>%
  summarise(
    n = n(),
    mean = mean(deaths),
    sd = sd(deaths),
    min = min(deaths),
    Q1 = quantile(deaths, 0.25),
    median = median(deaths),
    Q3 = quantile(deaths, 0.75),
    max = max(deaths)
  )
```

If we look at the mean, we might phrase our findings as: 'The average number of 
deaths in female-named hurricanes is about 24, compared to 14 in male named 
hurricanes.'. The reader could mentally translate this to: 'a typical 
female-named hurricane has 24 deaths, compared to 14 in male named hurricanes'.
But look at this:

```{r}
freq_abs <- with(subset(hur1, gender == "Female"), table(deaths <= 24))
freq_abs
```
and relative frequencies:
```{r}
prop.table(freq_abs)
```

Almost 80% of the female-named hurricanes actually heave a death toll no greater
than 24! Maybe '24' is not that typical after all.

What about the median? The median in the 2 groups is exactly identical, and it
tells us that in both female and male-named hurricanes, half have up to 5
deaths, and half have more:

```{r}
sort(subset(hur1, gender == "Female")$deaths)
```

```{r}
sort(subset(hur1, gender == "Male")$deaths)
```

This is also shown by the ECDF (kind of):
```{r}
ggplot(hur1, aes(x = deaths)) +
  stat_ecdf() +
  facet_grid(~ gender)
```

What about the influence of the year?
```{r}
ggplot(hur1, aes(x = year, y = deaths, color = gender)) +
  geom_point() +
  scale_color_colorblind()
```

Interesting! Up to 1970, names were almost exclusively from females. After 1970,
the distribution seems much more balanced:

```{r}
filter(hur1, year >= 1978) %>%
  ggplot(aes(x = year, y = deaths, color = gender)) +
  geom_point() +
  scale_color_colorblind()
```

How do the 2 group compare in this subset?
```{r}
filter(hur1, year >= 1978) %>%
  group_by(gender) %>%
  summarise(
    n = n(),
    mean = mean(deaths),
    sd = sd(deaths),
    min = min(deaths),
    Q1 = quantile(deaths, 0.25),
    median = median(deaths),
    Q3 = quantile(deaths, 0.75),
    max = max(deaths)
  )
```

In females, half of the hurricanes have between 3 and 11 deaths,
compared to 1-15 in males. The median is 5 in both groups.
