---
title: "R Notebook"
output: html_notebook
---

Set working directory to current one

```{r}
library(rstudioapi)
current_path = rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(current_path ))
```

Import Ble dataset
```{r}
Ble = read.table('Dataset_Ble.txt',header=TRUE,sep=';')
attach(Ble)
```
Initial check of imported data
```{r}
dim(Ble)
head(Ble)
```

One-way anova influence of variete on rdt
- Visual Inspection via boxplot
```{r}
boxplot(rdt~variete,col='yellow')
```
nb of observations by variete
```{r}
summary(variete)
```



Create model
```{r}
vLM = lm(rdt~variete)
```

Check assumptions on the gaussianity of the residuals

- Homogeneity of variances (leveneTest)
--> LeveneTest p-value small - there is no homogeneity of variances (what to do ???)
```{r}
if(!require(car)) install.packages("car")
library(car)
leveneTest(vLM)
```

- Gaussianity (QQplot)
--> QQPlot ok noise gaussian
```{r}
plot(vLM,2)
```
Shapiro
```{r}
shapiro.test(vLM)
```


Run Anova on model
--> p-value very small - there is an influence of variete on rdt
```{r}
AvLM = anova(vLM)
AvLM
```

One-way anova influence of phyto on rdt
- Visual Inspection via boxplot
```{r}
boxplot(rdt~phyto,col='red')
```
nb of observations by phyto
```{r}
summary(phyto)
```

Create model
```{r}
pLM = lm(rdt~phyto)
```

Check assumptions on the gaussianity of the residuals

- Homogeneity of variances (leveneTest)
--> LeveneTest p-value >0.05 - there is homogeneity of variances
```{r}
if(!require(car)) install.packages("car")
library(car)
leveneTest(pLM)
```

- Gaussianity (QQplot)
--> QQPlot ok noise gaussian
```{r}
plot(pLM,2)
```

Run Anova on model
--> p-value very big - there is no influence of phyto on rdt
```{r}
ApLM = anova(pLM)
ApLM
```

Two-way anova influence of phyto/variete on rdt
- Visual Inspection via boxplot

```{r}
boxplot(rdt ~ phyto * variete, col=c('green','blue'))
```

Create model
```{r}
vpLM = lm(rdt ~ phyto * variete)
```

Check assumptions on the gaussianity of the residuals

- Homogeneity of variances (leveneTest)
--> LeveneTest p-value >0.05 - there is homogeneity of variances
```{r}
if(!require(car)) install.packages("car")
library(car)
leveneTest(vpLM)
```

- Gaussianity (QQplot)
--> QQPlot ok noise gaussian
```{r}
plot(vpLM,2)
```

Run Anova on model
--> p-value phyto:variete very big - there is no influence of cross effect of phyto/variete on rdt
```{r}
AvpLM = anova(vpLM)
AvpLM
```

Use additive model instead
```{r}
ADDvpLM = lm(rdt ~ phyto + variete)
```
Homgeneity of variance + normality
```{r}
plot(ADDvpLM,1)
plot(ADDvpLM,2)
```


Anova Result

```{r}
ADDAvpLM = anova(ADDvpLM)
ADDAvpLM
```



TO-BE-CEHCKED

test - anova for unbalanced design (???)
```{r}
my_anova <- aov(rdt ~ phyto * variete)
Anova(my_anova, type = "III")
```



Others

Pairwise 
```{r}
vpAOV = aov(rdt ~ phyto * variete)
par(las=1)
plot(TukeyHSD(vpAOV))
```



